```{r}
library(tidyverse)
library(ChemmineR)
library(ChemmineOB)
library(scatterplot3d)
library(fmcsR)
```

### SMARTS substructure search of reactive functional groups
### Walters & Murcko (2002) (REOS SMARTS filter)

```{r}
medchem <- read.SDFset("/home/bioinfo/Documents/Somdeb/MedChem/sdfs/complete/medchem_all.sdf")
```

```{r}
validID <- validSDF(medchem)
```

```{r}
medvalid <- medchem[validID]
```

```{r}
class(medvalid)
```
```{r}
smartsSearchOB(medvalid, "S(=O)(=O)[F,Cl,Br,I]", uniqueMatches = FALSE)
```
```{r}
smartsSearchOB(medvalid, "[N;10,11;$(N(=O)~[O;H0;20,21])]", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "[HC]=O", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "[Cl,Br,I][CH2]", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "C1[O,N]C1", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "[#6]OP(=O)O[#6]", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "[CH2][CH2][CH2][CH2][CH2][CH2][CH2][CH2]", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "OO", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "C(=O)C(=O)", uniqueMatches = FALSE)
```

```{r}
smartsSearchOB(medvalid, "C(=O)[Cl,Br,I]", uniqueMatches = FALSE)
```
## ALL medchem molecules have passed the functional group filters
```{r}

```

## Scaffold hopped molecules

```{r}
scaf <- read.SDFset("/home/bioinfo/Documents/Somdeb/SAScore/SAScore/sascore_le_tricc.sdf")
```

```{r}
validid <- validSDF(scaf)
```

```{r}
scafvalid <- scaf[validid]
```

```{r}
class(scafvalid)
```

Faliures noted in c1 and i1

```{r}
a1 <- smartsSearchOB(scafvalid, "S(=O)(=O)[F,Cl,Br,I]", uniqueMatches = FALSE)
```
```{r}
b1 <- smartsSearchOB(scafvalid, "[N;10,11;$(N(=O)~[O;H0;20,21])]", uniqueMatches = FALSE)
```
```{r}
# There were 30 failures for this
c1 <- smartsSearchOB(scafvalid, "[HC]=O", uniqueMatches = FALSE)
```
```{r}
d1 <- smartsSearchOB(scafvalid, "[Cl,Br,I][CH2]", uniqueMatches = FALSE)
```
```{r}
e1 <- smartsSearchOB(scafvalid, "C1[O,N]C1", uniqueMatches = FALSE)
```
```{r}
f1 <- smartsSearchOB(scafvalid, "[#6]OP(=O)O[#6]", uniqueMatches = FALSE)
```
```{r}
g1 <- smartsSearchOB(scafvalid, "[CH2][CH2][CH2][CH2][CH2][CH2][CH2][CH2]", uniqueMatches = FALSE)
```
```{r}
h1 <- smartsSearchOB(scafvalid, "OO", uniqueMatches = FALSE)
```
```{r}
# There were 3 failures for this
i1 <- smartsSearchOB(scafvalid, "C(=O)C(=O)", uniqueMatches = FALSE)
```
```{r}
j1 <-smartsSearchOB(scafvalid, "C(=O)[Cl,Br,I]", uniqueMatches = FALSE)
```


#### Filter out compounds which failed SMARTS search
Use table(c1) or table(i1) to find out which ones

```{r}
c1_logical <- ifelse(c1=="0",TRUE,FALSE)
C1PASS <-scafvalid[c1_logical]
cid(C1PASS) <- makeUnique(cid(C1PASS))
```
```{r}
# C1PASS molecules (without aldehydes) which have passed the
# 1,2-Dicarbonyl group filter
i2 <- smartsSearchOB(C1PASS, "C(=O)C(=O)", uniqueMatches = FALSE)
i2_logical <- ifelse(i2=="0",TRUE,FALSE)
SCAFPASS <- C1PASS[i2_logical]
```

## Scaffold hopped molecules which have passed the Spark filters (1000->806),
## the synthetic accessibility < triacsinc (806->508), aldehyde filters
## (506->478), and 1,2-Dicarbonyl filters (478->476) = SCAFPASS

```{r}
write.SDF(SCAFPASS, file="/home/bioinfo/Documents/Somdeb/SAScore/SAScore/scaffoldhop_allfilterpass-476.sdf")
```

